\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{November 4}
\newcommand{\assignmentassignedate}{November 1}
\newcommand{\assignmentnumber}{Six}

\newcommand{\labyear}{2019}
\newcommand{\labday}{Friday}
\newcommand{\labdueday}{Monday}
\newcommand{\labtime}{9:00 am}

\newcommand{\assigneddate}{Assigned: \labday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labdueday, \assignmentduedate, \labyear{} at \labtime{}}

% Edit these commands to give the name to the main program

\newcommand{\mainprogram}{\lstinline{Experiment}}
\newcommand{\mainprogramsource}{\lstinline{src/main/java/practicalsix/experiment/Experiment.java}}

% Edit these commands to give the name to the test suite

\newcommand{\testprogram}{\lstinline{TestExperiment}}
\newcommand{\testprogramsource}{\lstinline{src/test/java/practicalsix/TestExperiment.java}}

% Edit this commands to describe key deliverables

\newcommand{\reflection}{\lstinline{writing/reflection.md}}

% Commands to describe key development tasks

% --> Running gatorgrader.sh
\newcommand{\gatorgraderstart}{\command{gradle grade}}
\newcommand{\gatorgradercheck}{\command{gradle grade}}

% --> Compiling and running and testing program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradletest}{\command{gradle test}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

% NOTE: Could be improved, problems due to nesting

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

\newcommand{\gitcommitmainprogram}{\command{git commit src/main/java/practicalsix/experiment/Experiment.java -m "Your
descriptive commit message"}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 101}
\lhead{Practical Assignment \assignmentnumber{}}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 101\\Data Abstraction\\
      Fall 2019\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Practical \assignmentnumber{} \\ \assigneddate{} \\ \duedate{}}

\section*{Objectives}

To continue to practice using GitHub to access the files for an assignment. You
will complete a programming project using the minimal source code provided by
the course instructor, ultimately implementing a framework for experimentally
evaluating two algorithms for arithmetic computation. Specifically, you will
create and test both an iterative and a recursive implementation of Java
programs that calculate numbers in the Factorial sequence. You will also
practice creating a data table and calculating an order of growth ratio, honing
your ability to analytically and empirically evaluate the worst-case time
complexity of an algorithm. Finally, you will continue to learn how to implement
and test a Java program, practicing the creation of Java classes on your own and
using an automated grading tool to assess your progress towards correctly
completing the project.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Follow each step carefully}. Slowly read each sentence in the
  assignment sheet, making sure that you precisely follow each instruction. Take
  notes about each step that you attempt, recording your questions and ideas and
  the challenges that you faced. If you are stuck, then please tell a technical
  leader or instructor what assignment step you recently completed.

\item {\bf Regularly ask and answer questions}. Please log into Slack at the
  start of a laboratory or practical session and then join the appropriate
  channel. If you have a question about one of the steps in an assignment, then
  you can post it to the designated channel. Or, you can ask a student sitting
  next to you or talk with a technical leader or the course instructor.

\item {\bf Store your files in GitHub}. Starting with this laboratory
  assignment, you will be responsible for storing all of your files (e.g., Java
  source code and Markdown-based writing) in a Git repository using GitHub
  Classroom. Please verify that you have saved your source code in your Git
  repository by using \command{git status} to ensure that everything is updated.
  You can see if your assignment submission meets the established correctness
  requirements by using the provided checking tools on your local computer and
  in checking the commits in GitHub.

\item {\bf Keep all of your files}. Don't delete your programs, output files,
  and written reports after you submit them through GitHub; you will need them
  again when you study for the quizzes and examinations and work on the other
  laboratory, practical, and final project assignments.

\item {\bf Explore teamwork and technologies}. While certain aspects of the
  laboratory assignments will be challenging for you, each part is designed to
  give you the opportunity to learn both fundamental concepts in the field of
  computer science and explore advanced technologies that are commonly employed
  at a wide variety of companies. To explore and develop new ideas, you should
  regularly communicate with your team and/or the student technical leaders.

\item {\bf Hone your technical writing skills}. Computer science assignments
  require to you write source code documentation and detailed notes of your
  experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct,
  remembering that computer scientists must effectively communicate and
  collaborate with their team members and the technical leaders and course
  instructor.

% \item {\bf Review the Honor Code on the syllabus}. While you may discuss your
%   assignments with others, copying source code or writing is a violation of
%   Allegheny College's Honor Code.

\end{itemize}

\section*{Reading Assignment}

If you have not done so already, please read all of the relevant ``GitHub
Guides'', available at \url{https://guides.github.com/}, that explain how to
use many of the features that GitHub provides. In particular, please make sure
that you have read guides such as ``Mastering Markdown'' and ``Documenting Your
Projects on GitHub''; each of them will help you to understand how to use both
GitHub and GitHub Classroom. To do well on this laboratory assignment, you
should also read Section 1.5.2 in the textbook, paying particularly close
attention to the material about iteration constructs. You should also review
Sections 4.1 through 4.3 and read Sections 5.1 through 5.5, focusing on Code
Fragment 5.13 and the text that describes and analyzes it. Please see the
course instructor or one of the technical leaders if you have questions about
these reading assignments.

\section*{Accessing the Practical Assignment on GitHub}

To access the laboratory assignment, you should go into the
\channel{\#announcements} channel in our Slack team and find the announcement
that provides a link for it. Copy this link and paste it into a web browser.
Now, you should accept the assignment and see that GitHub Classroom created a
new GitHub repository for you to access the assignment's starting materials and
to store the completed version of your assignment. Specifically, to access your
new GitHub repository for this assignment, please click the green ``Accept''
button and then click the link that is prefaced with the label ``Your assignment
has been created here''. If you accepted the assignment and correctly followed
these steps, you should have created a GitHub repository with a name like
``Allegheny-Computer-Science-101-Fall-2019/computer-science-101-fall-2019-practical-6-gkapfham''.
Remember, unless you provide the instructor with documentation of the
extenuating circumstances that you are facing, not accepting the assignment
means that you automatically receive a failing grade for it.

Before you move to the next step of this assignment, please make sure that you
read all of the content on the web site for your new GitHub repository, paying
close attention to the technical details about the commands that you will type
and the output that your program must produce. Now you are ready to download the
starting materials to your laboratory computer. Click the ``Clone or download''
button and, after ensuring that you have selected ``Clone with SSH'', please
copy this command to your clipboard. To enter into your course directory you
should now type \command{cd cs101F2019}. By typing \command{git clone} in your
terminal and then pasting in the string that you copied from the GitHub site you
will download all of the code for this assignment. For instance, if the course
instructor ran the \command{git clone} command in the terminal, it would look
like:

\begin{lstlisting}
  git clone git@github.com:Allegheny-Computer-Science-101-F2019/computer-science-101-fall-2019-practical-6-gkapfham.git
\end{lstlisting}

After this command finishes, you can use \command{cd} to change into the new
directory. If you want to \step{go back} one directory from your current
location, then you can type the command \command{cd ..}. Please continue to use
the \command{cd} and \command{ls} commands to explore the files that you
automatically downloaded from GitHub. What files and directories do you see?
What do you think is their purpose? Spend some time exploring, sharing your
discoveries with a neighbor and a \mbox{technical leaders}. Specifically, each
student should identify the Java methods that they must add to complete the
assignment and then understand how they connect to the provided classes. Can you
create a diagram showing all of class relationships? Make sure that you have the
instructor check your project diagram.

\section*{Evaluating the Performance of Recursion and Iteration}

This primary goal for this assignment is to afford students the opportunity to
practice creating their own Java classes and JUnit test cases. Critically, it is
important to note that the build of the provided source code will fail because
it is missing methods that you must implement. As such, it is your
responsibility to study the provided source code and add in all of the new code
that you need to ensure that the system compiles and runs an experiment to
evaluate the efficiency of the \program{RecursiveFactorialComputation} and the
\program{IterativeFactorialComputation}. Specifically, the equation for the
factorial sequence is defined as $n! = n \times n-1 \times n-2 \times \ldots
\times 1$ or $n! = n \times (n-1)!$ when $n \geq 1$, with $0! = 1$ defined as
the base case. Following this recursive equation, you can create and test an
implementation of the \program{public long compute(int value)} method in the
\program{RecursiveFactorialComputation} class. As in the previous assignments,
please make sure that you correctly set the name of this class so that it
integrates correctly into the experimentation framework. Finally, make sure that
you understand how this Java class relates to and is tested by the others that
are the needed for this practical assignment. Make sure to consult your diagram!

Using a loop, you should also create a \program{compute} method in
\program{IterativeFactorialComputation}. Students who want to learn more about
the numerical values in the factorial sequence are encouraged to visit the web
site \url{http://www.tsm-resources.com/alists/fact.html}. You can also learn
more about the properties of and values in the factorial sequence by visiting
the web site \url{https://www.mathsisfun.com/numbers/factorial.html}. Once you
have implemented methods that pass your own test suite and compute the correct
numerical values, you are ready to run a doubling experiment. Please proceed
cautiously as you run the experiments: it is possible that one or more of the
algorithms may not be efficient, thus potentially leading to either extended
periods of time to run an experiment campaign or program crashes. With that
said, if it is necessary to do so you should determine what inputs lead to poor
performance and/or crashes and then explain why they are evident. Finally,
please bear in mind that the numerical values in the factorial sequence grow
very rapidly and thus even the \program{long} data type may not be able to
correctly store the computed values.

Figure~\ref{fig:code} provides a correct implementation of a \program{compute}
method that recursively calculates a requested number in the factorial sequence.
%
Although you do not need to exactly adopt this solution, you should carefully
study it to ensure that you can identify the base case(s) and the recursive
case. Finally, please make sure that you can classify this method as exhibiting
linear, binary, or multiple recursion.
%
After reviewing this method's source code, do you have an intuition about how
its performace would compare to the recursive implementation for the Fibonacci
sequence?

\begin{figure}[t]
  \centering

  \begin{verbatim}
  public long compute(int chosenValue) {
    long result = 1L;
    if ( chosenValue == 0 ) {
      return result;
    } else if ( chosenValue == 1 ) {
      return result;
    } else {
      return chosenValue * compute(chosenValue - 1);
    }
  }
  \end{verbatim}

  \vspace*{-.5in}
  \caption{A Suggestion for How to Recursively Compute the Factorial Number.}~\label{fig:code}
  \vspace*{-.25in}
\end{figure}


\section*{Checking the Correctness of Your Program}

As verified by Checkstyle tool, the source code for \mainprogram{} and its test
suite must adhere to all of the requirements in the Google Java Style Guide
available at \url{https://google.github.io/styleguide/javaguide.html}. Instead
of requiring you to manually check that your deliverables adhere to these
industry-accepted standards, tools will make it easy for you to automatically
check if your submission meets the standards for correctness. For instance,
Gradle will run the provided tests, while GatorGrader will check to ensure that
\mainprogram{} uses the correct number of \command{println} statements and that
you committed to GitHub a sufficient number of times during this assignment. To
get started with the use of GatorGrader, type the command \gatorgraderstart{}
in your Docker container. If you have mistakes in your assignment, then you will
need to review GatorGrader's output, find the mistake, and try to fix it. Once
your program is building correctly, fulfilling at least some of the
assignment's requirements, you should transfer your files to GitHub using the
\gitcommit{} and \gitpush{} commands. For example, if you want to signal that
the \mainprogramsource{} file has been changed and is ready for transfer to
GitHub you would first type \gitcommitmainprogram{} in your terminal, followed
by typing \gitpush{}.

After the course instructor enables \step{continuous integration} with a system
called Travis CI, when you use the \gitpush{} command to transfer your source
code to your GitHub repository, Travis CI will initialize a \step{build} of your
assignment, checking to see if it meets all of the requirements. If both your
source code and documentation meet all of the established requirements, then you
will see a green \checkmark{} in the listing of commits in GitHub after awhile.
If your submission does not meet the requirements, a red \naughtmark{} will
appear instead. The instructor will reduce a student's grade for this assignment
if the red \naughtmark{} appears on the last commit in GitHub immediately before
the assignment's due date. Yet, if the green \checkmark{} appears on the last
commit in your GitHub repository, then you satisfied all of the main checks,
thereby allowing the course instructor to evaluate other aspects of your source
code and writing, as further described in the \step{Evaluation} section of this
assignment sheet. Unless you provide the instructor with documentation of the
extenuating circumstances that you are facing, no late work will be considered
towards your grade for this practical assignment.

\section*{Summary of the Required Deliverables}

\noindent Students do not need to submit printed source code or technical
writing for any assignment in this course. Instead, this assignment invites you
to submit, using GitHub, the following deliverables.

\begin{enumerate}

  \setlength{\itemsep}{0in}

\item A properly documented and correct version of all the Java source files
  (i.e., factorial methods, experiments, and tests) that meets the set
  requirements and produces the desired output.
  %
  Although you are not required to complete a technical writing assignment as
  part of this project, you should make sure that you understand and can discuss
  the resulting data tables.

\end{enumerate}

\section*{Evaluation of Your Practical Assignment}

Using a report that the instructor shares with you through the commit log in
GitHub, you will privately received a grade on this assignment and feedback on
your submitted deliverables. Your grade for the assignment will be a function of
the whether or not it was submitted in a timely fashion and if your program
received a green \checkmark{} indicating that it met all of the baseline
requirements checked by GatorGrader.
%
Please remember to read your GitHub repository's \program{README.md} file for a
description of the completion grade that you will receive for this practical
assignment.
%
You can talk with the instructor if you have questions about the evaluation of
this practical assignment.

\section*{Adhering to the Honor Code}

In adherence to the Honor Code, students should complete this assignment on an
individual basis. While it is appropriate for students in this class to have
high-level conversations about the assignment, it is necessary to distinguish
carefully between the student who discusses the principles underlying a problem
with others and the student who produces assignments that are identical to, or
merely variations on, someone else's work. Deliverables (e.g., Java source code
or Markdown-based technical writing) that are nearly identical to the work of
others will be taken as evidence of violating the \mbox{Honor Code}. Please see
the course instructor if you have questions about this policy.

\end{document}
