\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{September 2}
\newcommand{\assignmentassignedate}{August 30}
\newcommand{\assignmentnumber}{One}

\newcommand{\labyear}{2019}
\newcommand{\labdueday}{Monday}
\newcommand{\labassignday}{Friday}
\newcommand{\labtime}{9:00 am}

\newcommand{\assigneddate}{Assigned: \labassignday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labdueday, \assignmentduedate, \labyear{} at \labtime{}}

% Edit these commands to give the name to the main program

\newcommand{\mainprogram}{\lstinline{Swap}}
\newcommand{\mainprogramsource}{\lstinline{src/main/java/practicalone/Swap.java}}

% Edit these commands to give the name to the test suite

\newcommand{\testprogram}{\lstinline{TestSwap}}
\newcommand{\testprogramsource}{\lstinline{src/test/java/practicalone/TestSwap.java}}

% Commands to describe key development tasks

% --> Running gatorgrader.sh
\newcommand{\gatorgraderstart}{\command{gradle grade}}
\newcommand{\gatorgradercheck}{\command{gradle grade}}

% --> Compiling and running and testing program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradletest}{\command{gradle test}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

% NOTE: Could be improved, problems due to nesting

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

\newcommand{\gitcommitmainprogram}{\command{git commit src/main/java/practicalone/Swap.java -m "Your
descriptive commit message"}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 101}
\lhead{Practical Assignment \assignmentnumber{}}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 101\\Data Abstraction\\
      Fall 2019\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Practical \assignmentnumber{} \\ \assigneddate{} \\ \duedate{}}

% Slack for this course:

% https://join.slack.com/t/cmpsc101fall2019/signup

\section*{Objectives}

To learn how to install and use a contained-based platform, called Docker, to
support the completion of technical activities (e.g., building and running a
Java program) during the class, laboratory, and practical sessions. In addition
to learning how to install and use a text editor such as Atom or VS Code,
student will also learn how to install a package manager such as Brew or
Chocolatey. After you install a package manager, you will use it to further
install programs like a terminal window and the Git command-line client. As
necessary, students may optionally configure their laptops and GitHub accounts
--- although a later assignment will also review the steps needed to complete
this task. Throughout this assignment, students will also learn how to use Slack
to support communication with each other, the student technical leaders, and the
course instructor.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Follow each step carefully}. Slowly read each sentence in the
  assignment sheet, making sure that you precisely follow each instruction. Take
  notes about each step that you attempt, recording your questions and ideas and
  the challenges that you faced. If you are stuck, then please tell a technical
  leader or instructor what assignment step you recently completed.

\item {\bf Regularly ask and answer questions}. Please log into Slack at the
  start of a laboratory or practical session and then join the appropriate
  channel. If you have a question about one of the steps in an assignment, then
  you can post it to the designated channel. Or, you can ask a student sitting
  next to you or talk with a technical leader or the course instructor.

% \item {\bf Store your files in GitHub}. Starting with this practical assignment,
%   you will be responsible for storing all of your files (e.g., Java source code
%   and Markdown-based writing) in a Git repository using GitHub Classroom. Please
%   verify that you have saved your source code in your Git repository by using
%   \command{git status} to ensure that everything is updated. You can see if your
%   assignment submission meets the established correctness requirements by using
%   the provided checking tools on your local computer and in checking the commits
%   in GitHub.

% \item {\bf Keep all of your files}. Don't delete your programs, output files,
%   and written reports after you submit them through GitHub; you will need them
%   again when you study for the quizzes and examinations and work on the other
%   laboratory, practical, and final project assignments.

% \item {\bf Back up your files regularly}. All of your files are regularly
%   backed-up to the servers in the Department of Computer Science and, if you
%   commit your files regularly, stored on GitHub. However, you may want to use a
%   flash drive, Google Drive, or your favorite backup method to keep an extra
%   copy of your files on reserve. In the event of any type of system failure, you
%   are responsible for ensuring that you have access to a recent backup copy of
%   all your files.

\item {\bf Explore teamwork and technologies}. While certain aspects of the
  course assignments will be challenging for you, each part is designed to give
  you the opportunity to learn both fundamental concepts in the field of
  computer science and explore advanced technologies that are commonly employed
  at a wide variety of companies. To explore and develop new ideas, you should
  regularly communicate with your team and/or the student technical leaders.

\item {\bf Hone your technical writing skills}. Computer science assignments
  require to you write technical documentation and descriptions of your
  experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct,
  remembering that computer scientists must effectively communicate and
  collaborate with their team members, the student technical leaders, and the
  course instructor.

\item {\bf Review the Honor Code on the syllabus}. While you may discuss your
  assignments with others, copying source code or writing is a violation of
  Allegheny College's Honor Code.

\end{itemize}

\section*{Reading Assignment}

If you have not done so already, please read all of the relevant ``GitHub
Guides'', available at \url{https://guides.github.com/}, that explain how to use
many of the features that GitHub provides. In particular, please make sure that
you have read guides such as ``Mastering Markdown'' and ``Documenting Your
Projects on GitHub''; each of them will help you to understand how to use both
GitHub and GitHub Classroom. In addition to reading all of the relevant
documentation about Docker that is available at \url{https://docs.docker.com/},
students should review the documentation about their chosen text editor (e.g.,
Vim, Atom, or VS Code) and package manager (e.g., Chocolatey or Brew); to
complete this part of the reading assignment, please visit the web site of each
tool.

Once you have bookmarked references for each of the aforementioned web sites and
reviewed some of the printed guides, please access the Slack workspace that we
will use in this course, available for signup at
\url{https://join.slack.com/t/cmpsc101fall2019/signup}. Once you have access to
the Slack workspace, please join the \url{#practicals} channel and post one
thing that you have learned about Docker, your chosen text editor, and your
package manager. In addition to posting what you learned, please review the
comments from other students, always attempting to post something that is not
already in the channel. Please make sure that you keep Slack open throughout
your completion of this practical assignment as you will continue to use it to
communicate with other students, the technical leaders, and the course
instructor, who will read and assess your insights.

\section*{Installing and Using A Text Editor}

You should first install a text editor, such as Atom or VS Code, which are
respectively available from \url{https://atom.io/} and
\url{https://code.visualstudio.com/}. Please follow the download instructions
for your operating system to install one or both of these text editors.
Optionally, you may also explore the installation of other editors such as Vim.
Students who want to install Vim or Neovim should talk with the course
instructor or a student technical leader. After choosing a text editor, return
to the \url{#practicals} channel and say which editor you picked and why you
selected it.

\section*{Installing and Using a Package Manager}

If you are using the Linux operating system, then you can use the default
\program{apt} program to install other programs. Since Windows and MacOS do not
provide a default package manager, you should install \program{choco} for
Windows and \program{brew} for MacOS. To learn more about these tools, you can
respectively visit \url{https://chocolatey.org/} and \url{https://brew.sh/}.
Please follow the installation instructions for these package managers, making
note of the challenges that you faced and the ways in which you solved them.
Finally, please share your experiences in the correct Slack channel.

After installing the package manager, you can use it to install other programs!
For example, the Windows operating system does not have a good terminal window
by default. To handle this situation, you could install the new Windows Terminal
by using the Chocolatey commands at
\url{https://chocolatey.org/packages/microsoft-windows-terminal}. Since Linux
and MacOS have suitable terminal windows available by default, everyone can now
use a terminal to run package manager commands to install the \program{git}
command-line tool for their operating system. Before moving onto the next step
of the assignment, you should now have working versions of the following
software tools:

\begin{itemize}

  \item Text editor

  \item Package manager

  \item Terminal window

  \item Git command-line client

\end{itemize}

\section*{Installing and Using Docker}

Docker is a platform for software engineers and system administrators to
develop, deploy, and run software applications with containers that house
full-fledged executable programs. When software engineers use Docker, they no
longer have to focus on installing the programs that come inside of the Docker
container. Instead, they can use a program in a Docker container as if it was
installed on their development workstation, regardless of whether they are using
Windows, Linux, or MacOS. A wide variety of companies, such as Visa and PayPal,
use Docker. Throughout this semester, you will receive Docker containers that
house programs like Java, Gradle, and GatorGrader. These containers will also
hold the supporting libraries and packages needed to run the aforementioned
programs. With that said, please note that you should not store source code or
technical writing inside of a Docker container because it is a ``transient''
software application that does not store its state upon shut-down. Finally, the
provided Docker container will not include programs like Git or your text editor
as the instructor expects that each student will install these programs
separately.

The Department of Computer Science has adopted Docker for all computer science
courses. The course instructor, student technical leaders, and the departmental
systems administrator will support the use of Docker on laptops running Windows,
Linux, and MacOS. The instructor in this course will provide Docker containers
with all of the necessary software to run all class exercises, laboratory, and
practical assignments. Students will then be able to develop and run Java
programs on their own laptops, using an instructor-provided Docker container.
Before being able to use Docker, students must first complete the installation
of it, following steps that are specific to the operating system on your laptop.
Bearing in mind that you must follow steps customized to your laptop, please
talk to the instructor or a technical leader if you get stuck on these steps.

Before reading these installation instructions for Docker, you should know that
they are largely customized for students who use Windows or MacOS. If you use
the Linux operating system, then you can follow the installation instructions
available at this web site:
\url{https://docs.docker.com/install/linux/docker-ce/ubuntu/}. Okay, now you are
ready to install and use Docker!

\begin{figure}

\begin{verbatim}
docker run hello-world

docker : Unable to find image 'hello-world:latest' locally
...

latest:
Pulling from library/hello-world
ca4f61b1923c:
Pulling fs layer
ca4f61b1923c:
Download complete
ca4f61b1923c:
Pull complete
Digest: sha256:97ce6fa4b6cdc0790cda65fe7290b74cfebd9fa0c9b8c38e979330d547d22ce1
Status: Downloaded newer image for hello-world:latest

Hello from Docker!
This message shows that your installation appears to be working correctly.
...
\end{verbatim}

\vspace*{-.25in}
\caption{Example Output from Running a ``Hello World!''  Docker Container.}~\label{fig:docker}
\vspace*{-.25in}
\end{figure}

\begin{enumerate}

    \item Determine the version of your operating system. If your laptop meets
      the requirements outlined below, then you can proceed to the next step.
      For older MacOS and Windows operating systems, including Windows 10 Home,
      please read the installation instructions for Docker Toolbox, provided at
      \url{https://docs.docker.com/toolbox/toolbox_install_windows/}.

      \textbf{Docker Desktop Specifications} \\
      -- Mac:
      \begin{itemize}
        \item 2010 model or newer with hardware support for MMU, EPT, and Unrestricted Mode
        \item MacOS Sierra 10.12 or newer
        \item 4 GB of RAM
        \item VirtualBox prior to version 4.3.30 cannot be installed
      \end{itemize}
      -- Windows:
      \begin{itemize}
        \item Windows 10 64bit Pro, Enterprise, or Education (Build 15063 or later)
        \item Virtualization is enabled in BIOS
        \item CPU SLAT-capable feature
        \item 4 GB of RAM
      \end{itemize}
      -- Linux:
      \begin{itemize}
        \item 64bit
        \item Kernel 3.10 or later
      \end{itemize}

    \item Go to \url{https://docs.docker.com/install/} and  from the menu on the
      left-hand side select and follow the installation tutorial for your
      operating system (Linux, MacOS, or Windows). For example, if you have
      MacOS you can find the installation guide at
      \url{https://docs.docker.com/docker-for-mac/install/}. If you have the
      Windows operating system, please follow the installation guide at
      \url{https://docs.docker.com/docker-for-windows/install/}. Students with
      Linux can use a package manager when following the instructions at the
      aforementioned site.

      Please note that the Docker Desktop for Mac and Windows requires you to
      create an account to download the platform. To bypass this, you can use
      the following direct download links:

    \begin{itemize}
        \item Windows: \\ \url{https://download.docker.com/win/stable/Docker\%20for\%20Windows\%20Installer.exe}
        \item Mac: \url{https://download.docker.com/mac/stable/Docker.dmg}
    \end{itemize}

  \item Once your Docker installation is complete, please run a terminal window
    on your machine. In MacOS and Linux, you can search for a ``terminal'' and
    on Windows machines look for \command{cmd}. If you correctly followed the
    previous assignment steps and you use Windows, then you can also use the new
    ``terminal'' that you installed. Now, in the terminal window you can type
    \command{docker run hello-world} command and press ``Enter''. If you see an
    output similar to the one in Figure~\ref{fig:docker} you can high five a
    technical leader as your Docker set up was successful!

  \item Note that even though the previous Docker container caused your terminal
    to print a diagnostic message, it did not allow you to run any programs.
    With that said, it is also possible to download and use a Docker container
    that contains the complete version of an operating system! To explore this
    feature of Docker, please type the command \command{docker run -it ubuntu
    /bin/bash} and notice that it produces output like that found in
    Figure~\ref{fig:ubuntu}. Congratulations, you are now running the Linux
    operating system inside a container running on your host operating system!
    If you got stuck on this step, please ask the course instructor for help.

  \item You can use commands line \command{cd} and \command{ls} to explore the
    default configuration of this Ubuntu container provided by Docker. Outside
    of Docker, type the commands \command{docker ps} and \command{docker images}
    and explain this output to a technical leader and a fellow student. You
    should also paste the output of these commands into the \url{#practicals}
    channel in Slack. Remember, you can paste a block of formatted text into
    Slack by using three ``backticks'', pasting the content, and then closing
    with three additional backticks. Finally, please write in Slack about what
    you experienced during the installation of Docker. What did you find
    confusing? Why did you find it to be confusing? Based on your experiences in
    this practical assignment, what do you think are the benefits associated
    with using Docker in a computer science course?

\end{enumerate}


\begin{figure}

\begin{verbatim}
Unable to find image 'ubuntu:latest' locally
latest: Pulling from library/ubuntu
35c102085707: Pull complete
251f5509d51d: Pull complete
8e829fe70a46: Pull complete
6001e1789921: Pull complete
Digest: sha256:d1d454df0f579c6be4d8161d227462d69e163a8ff9d20a847533989cf0c94d90
Status: Downloaded newer image for ubuntu:latest
\end{verbatim}

\vspace*{-.25in}
\caption{Example Output from Running a Ubuntu Docker Container.}~\label{fig:ubuntu}
\vspace*{-.25in}
\end{figure}

\vspace*{-.05in}

After completing each stage of this practical assignment, you should have these
correctly working versions of the following program installed on your laptop,
regardless of your operating system.

\vspace*{-.05in}

\begin{itemize}
  \setlength{\itemsep}{0pt}

  \item Text editor

  \item Package manager

  \item Terminal window

  \item Git command-line client

  \item Docker container platform with:

    \vspace*{-.1in}
    \begin{itemize}
      \setlength{\itemsep}{0pt}

      \item Docker ``Hello World!'' container
      \item Docker Ubuntu container

    \end{itemize}

\end{itemize}

\vspace*{-.1in}

Every student in this course is must have access to a Windows, MacOS, or Linux
laptop that correctly runs these programs, in full support of forthcoming Java
software development projects.

\section*{Summary of the Required Deliverables}

% \noindent Students do not need to submit printed source code or technical
% writing for any assignment in this course.

\noindent This assignment invites you to submit, using \url{#practicals} in
Slack, the following deliverables:

\vspace*{-.025in}

\begin{enumerate}

  \setlength{\itemsep}{0in}

  \item A description of the challenges that you experienced and the solutions
    that you developed for each challenge, following each of the prompts
    previously given in the assignment sheet.

  \item The correctly formatted and fully explained output from running Docker
    commands like \command{docker ps} and \command{docker images} in your
    correctly configured terminal window.

\end{enumerate}

\vspace*{-.075in}

\section*{Evaluation of Your Practical Assignment}

All practical assignments in this course are graded on a mastery, or
credit-no-credit, basis. For this assignment, you will receive credit as long as
you type all of the required information into the \url{#practicals} channel in
Slack. You will know that you earned the checkmark grade for this assignment if
you can see a checkmark emoji next to a comment that you made in the appropriate
channel. Please bear in mind that, for future practical and laboratory
assignments, you will submit all of your source code and technical writing
through a GitHub repository created by GitHub Classroom. However, for this
assignment, you must instead submit all of your work through Slack.

\end{document}
