\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{September 23}
\newcommand{\assignmentassignedate}{September 20}
\newcommand{\assignmentnumber}{Two}

\newcommand{\labyear}{2019}
\newcommand{\labdueday}{Monday}
\newcommand{\labassignday}{Friday}
\newcommand{\labtime}{9:00 am}

\newcommand{\assigneddate}{Assigned: \labassignday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labdueday, \assignmentduedate, \labyear{} at \labtime{}}

% Edit these commands to give the name to the main program

\newcommand{\mainprogram}{\lstinline{StatisticsCalculator}}
\newcommand{\mainprogramsource}{\lstinline{src/main/java/practicaltwo/StatisticsCalculator.java}}

% Edit these commands to give the name to the test suite

\newcommand{\testprogram}{\lstinline{TestStatisticsCalculator}}
\newcommand{\testprogramsource}{\lstinline{src/test/java/practicaltwo/TestStatisticsCalculator.java}}

% Edit this commands to describe key deliverables

\newcommand{\reflection}{\lstinline{writing/reflection.md}}

% Commands to describe key development tasks

% --> Running gatorgrader.sh
\newcommand{\gatorgraderstart}{\command{gradle grade}}
\newcommand{\gatorgradercheck}{\command{gradle grade}}

% --> Compiling and running and testing program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradletest}{\command{gradle test}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

% NOTE: Could be improved, problems due to nesting

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

\newcommand{\gitcommitmainprogram}{\command{git commit src/main/java/practicaltwo/StatisticsCalculator.java -m "Your
descriptive commit message"}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 101}
\lhead{Practical Assignment \assignmentnumber{}}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 101\\Data Abstraction\\
      Fall 2019\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Practical \assignmentnumber{} \\ \assigneddate{} \\ \duedate{}}

\section*{Objectives}

To continue to practice using GitHub to access the files for a course
assignment. Additionally, to review how to use the Ubuntu operating system and
software development programs such as a ``terminal window'' and the ``Docker
desktop''. You will also continue to practice using Slack to support
communication with the technical leaders and the instructor. Next, you will
learn how to implement and test a Java program that uses arrays and loops to
compute summary statistics for a list of data points, also practicing how to use
an automated grading tool to assess your progress.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Follow each step carefully}. Slowly read each sentence in the
  assignment sheet, making sure that you precisely follow each instruction. Take
  notes about each step that you attempt, recording your questions and ideas and
  the challenges that you faced. If you are stuck, then please tell a technical
  leader or instructor what assignment step you recently completed.

\item {\bf Regularly ask and answer questions}. Please log into Slack at the
  start of a laboratory or practical session and then join the appropriate
  channel. If you have a question about one of the steps in an assignment, then
  you can post it to the designated channel. Or, you can ask a student sitting
  next to you or talk with a technical leader or the course instructor.

\item {\bf Store your files in GitHub}. Starting with this laboratory
  assignment, you will be responsible for storing all of your files (e.g., Java
  source code and Markdown-based writing) in a Git repository using GitHub
  Classroom. Please verify that you have saved your source code in your Git
  repository by using \command{git status} to ensure that everything is updated.
  You can see if your assignment submission meets the established correctness
  requirements by using the provided checking tools on your local computer and
  in checking the commits in GitHub.

\item {\bf Keep all of your files}. Don't delete your programs, output files,
  and written reports after you submit them through GitHub; you will need them
  again when you study for the quizzes and examinations and work on the other
  laboratory, practical, and final project assignments.

\item {\bf Explore teamwork and technologies}. While certain aspects of the
  laboratory assignments will be challenging for you, each part is designed to
  give you the opportunity to learn both fundamental concepts in the field of
  computer science and explore advanced technologies that are commonly employed
  at a wide variety of companies. To explore and develop new ideas, you should
  regularly communicate with your team and/or the student technical leaders.

\item {\bf Hone your technical writing skills}. Computer science assignments
  require to you write source code documentation and detailed notes of your
  experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct,
  remembering that computer scientists must effectively communicate and
  collaborate with their team members and the technical leaders and course
  instructor.

\item {\bf Review the Honor Code on the syllabus}. While you may discuss your
  assignments with others, copying source code or writing is a violation of
  Allegheny College's Honor Code.

\end{itemize}

\section*{Reading Assignment}

If you have not done so already, please read all of the relevant ``GitHub
Guides'', available at \url{https://guides.github.com/}, that explain how to use
many of the features that GitHub provides. In particular, please make sure that
you have read guides such as ``Mastering Markdown'' and ``Documenting Your
Projects on GitHub''; each of them will help you to understand how to use both
GitHub and GitHub Classroom. To do well on this assignment, you should also read
Chapters 1 and 2 in the textbook, paying particularly close attention to
Sections 1.3 through 1.9. Please see the instructor or a teaching assistant if
you have questions on these sections.

\section*{Accessing the Practical Assignment on GitHub}

To access the laboratory assignment, you should go into the
\channel{\#announcements} channel in our Slack team and find the announcement
that provides a link for it. Copy this link and paste it into a web browser.
Now, you should accept the assignment and see that GitHub Classroom created a
new GitHub repository for you to access the assignment's starting materials and
to store the completed version of your assignment. Specifically, to access your
new GitHub repository for this assignment, please click the green ``Accept''
button and then click the link that is prefaced with the label ``Your
assignment has been created here''. If you accepted the assignment and
correctly followed these steps, you should have created a GitHub repository
with a name like
``Allegheny-Computer-Science-101-Fall-2019/computer-science-101-fall-2019-practical-2-gkapfham''.
Unless you provide the instructor with documentation of the extenuating
circumstances that you are facing, not accepting this practical assignment
means that you automatically receive a failing grade for it.

Before you move to the next step of this assignment, please make sure that you
read all of the content on the web site for your new GitHub repository, paying
close attention to the technical details about the commands that you will type
and the output that your program must produce. Now you are ready to download the
starting materials to your laboratory computer. Click the ``Clone or download''
button and, after ensuring that you have selected ``Clone with SSH'', please
copy this command to your clipboard. At this point, you can open a new terminal
window and type the command \command{cd cs101F2019} to enter the course
directory. Now type \command{git clone} in your terminal and paste in the string
that you copied from the GitHub site to download this assignment's code. For
instance, if the instructor ran the \command{git clone} command, it would look
like:

\begin{lstlisting}
  git clone git@github.com:Allegheny-Computer-Science-101-F2019/computer-science-101-fall-2019-practical-2-gkapfham.git
\end{lstlisting}

After this command finishes, you can use \command{cd} to change into the new
directory. If you want to \step{go back} one directory from your current
location, then you can type the command \command{cd ..}. Please continue to use
the \command{cd} and \command{ls} commands to explore the files that you
automatically downloaded from GitHub. What files and directories do you see?
What do you think is their purpose? Spend some time exploring, sharing your
discoveries with a neighbor and a \mbox{student technical leader}.

\section*{Enhancing a Java Program that Calculates Summary Statistics}

Please study the source code of the \mainprogram, noticing that some of its
methods is incomplete. For instance, you will see that the program is missing an
implementation of the \program{calculateVariance} method. Can you add in a
correct version of this method? To learn more about how to perform a summation
to calculate the variance of a list of \program{Double} values, please read the
online article available at \url{https://www.wikihow.com/Calculate-Variance}.
Specifically, the article's equation asks you to iterate through all of the
values in an array so that you can subtract the mean from each data point and
then square the result. You will need to ensure that the result from each
iteration carries over to the next one. Then, you will perform a division and
return this value. Once you have finished this step, please do a ``project
wide'' search for the \command{TODO} markers inside of the provided Java source
code. Make sure that you complete each of the requested tasks as you are working
on this assignment. If you are stuck, please ask the course instructor for help!

You should run the test suite called \testprogram{} that is written in the JUnit
testing framework. Please read each of the tests as they explain, in Java source
code, the expected properties of the methods for calculating summary statistics.
If you want to \step{build} your program you can type the command \gradlebuild{}
in the Docker container, thereby causing the Java compiler to check your program
for errors and get it ready to run. If you see any error messages, go back to
your text editor and figure out what you mis-typed and fix it. Once you have
solved the problem, make a note of the error and the solution for resolving it.
Re-save your program and then build it again by re-running the \gradlebuild{}.
If you cannot build \mainprogram{} correctly, then please talk with a technical
leader or the instructor. You can also run the test suite for the \mainprogram{}
by typing \gradletest{} in the terminal. What is the purpose of each of these
test cases? How do they check the output of the methods that calculate summary
statistics? You should revise your implementation of \command{calculateVariance}
until all the tests pass.

After you resolve all of the building and testing errors, you can run your
program by typing \gradlerun{} in the terminal window---this is the ``execute''
step that will run your program and produce the designated output. Do you see
that it produces the correct output with the expected numerical values? For
instance, you program should output \command{This is the variance: 23.04}. If
not, then please determine what is wrong and try to repair the program and
re-build and re-run it.
%
Overall, please make sure that you have provide an implementation of all the
required methods in \mainprogram{} and that, when run in Docker, it produces the
output shown in Figure~\ref{fig:output}.

\begin{figure}[t]
  \centering

  \begin{verbatim}
    Displaying the values in the array.
      12.5
      6.2
      1.5
      2.9
      10.8
    Displaying the output with two decimal places.
    This is the arithmetic mean: 6.78
    This is the variance: 23.04
    This is the standard deviation: 4.80
  \end{verbatim}

  \vspace*{-.25in}
  \caption{The Expected Output of the \mainprogram{}.}~\label{fig:output}
  \vspace*{-.25in}

\end{figure}

\section*{Checking the Correctness of Your Program}

As verified Checkstyle, the Java source code must adhere to all of the
requirements in the Google Java Style Guide available at
\url{https://google.github.io/styleguide/javaguide.html}. The Markdown file that
contains your reflection must adhere to the standards described in the Markdown
Syntax Guide \url{https://guides.github.com/features/mastering-markdown/}.
Finally, your \reflection{} file should adhere to the Markdown standards
established by the \step{Markdown linting} tool available at
\url{https://github.com/markdownlint/markdownlint/} and the writing standards
set by the \step{prose linting} tool from \url{http://proselint.com/}. Instead
of requiring you to manually check that your deliverables adhere to these
industry-accepted standards, Gradle and GatorGrader make it easy for you to
automatically check your submission. For instance, Gradle will run the provided
JUnit tests, check to ensure that \mainprogram{} produces exactly ten lines of
output, and GatorGrader will see if you committed to GitHub enough times when
completing this assignment.

To get started with the use of GatorGrader, type the command \gatorgraderstart{}
in your Docker container. If you have mistakes in your assignment, then you will
need to review GatorGrader's output, find the mistake, and try to fix it. Once
your program is building correctly, fulfilling at least some of the assignment's
requirements, you should transfer your files to GitHub using the \gitcommit{}
and \gitpush{} commands. For example, if you want to signal that the
\mainprogramsource{} file has been changed and is ready for transfer to GitHub
you would first type \gitcommitmainprogram{} in your terminal, followed by
typing \gitpush{}.

After the course instructor enables \step{continuous integration} with a system
called Travis CI, when you use the \gitpush{} command to transfer your source
code to your GitHub repository, Travis CI will initialize a \step{build} of
your assignment, checking to see if it meets all of the requirements. If both
your source code and writing meet all of the established requirements, then you
will see a green \checkmark{} in the listing of commits in GitHub after awhile.
If your submission does not meet the requirements, a red \naughtmark{} will
appear instead. The instructor will reduce a student's grade for this
assignment if the red \naughtmark{} appears on the last commit in GitHub
immediately before the assignment's due date. Yet, if the green \checkmark{}
appears on the last commit in your GitHub repository, then you satisfied all of
the main checks, thereby allowing the course instructor to evaluate other
aspects of your source code and writing, as further described in the
\step{Evaluation} section of this assignment sheet. Unless you provide the
instructor with documentation of the extenuating circumstances that you are
facing, no late work will be considered towards your grade for this practical
assignment.

Please note that if you are completing this practical assignment on your own
laptop the GatorGrader program will only run correctly if you have installed
all of the programs on which it depends. You can see these dependencies (i.e.,
Gradle, Java, Markdownlint, Proselint, and Python) by reading the documentation
in your GitHub repository for this assignment. Note that assignment checking
will still work with Travis CI even if you do not have these programs installed
on your laptop. Please see the instructor or a technical leader if you have
trouble installing these programs.

\section*{Summary of the Required Deliverables}

\noindent Students do not need to submit printed source code or technical
writing for any assignment in this course. Instead, this assignment invites you
to submit, using GitHub, the following deliverables.

\vspace*{-.05in}
\begin{enumerate}

  \setlength{\itemsep}{0in}

\item A properly documented and correct version of \mainprogramsource{} that
  meets the established requirements and produces the desired output.

\end{enumerate}
\vspace*{-.1in}

\section*{Evaluation of Your Practical Assignment}

Using a report that the instructor shares with you through the commit log in
GitHub, you will privately received a grade on this assignment and feedback on
your submitted deliverables. Your grade for the assignment will be a function of
the whether or not it was submitted in a timely fashion and if your program
received a green \checkmark{} indicating that it met all of the baseline
requirements checked by GatorGrader.
%
Please remember to read your GitHub repository's \program{README.md} file for a
description of the completion grade that you will receive for this laboratory
assignment.
%
You can talk with the instructor if you have questions about the evaluation of
this practical assignment.

In adherence to the Honor Code, students should complete this assignment on an
individual basis. While it is appropriate for students in this class to have
high-level conversations about the assignment (concerning, for instance, the
equation for the variance summary statistic), it is necessary to distinguish
carefully between the student who discusses the principles underlying a
technical problem with others and the student who produces assignment
deliverables (e.g., source code) that are identical to, or merely variations on,
someone else's work. Please talk with the course instructor if you are not sure
how to abide by the Honor Code while completing this assignment.

\end{document}
